---
output:
  pdf_document:
    keep_tex: yes
  html_document: default
---
#Title: Statistical Inference Course Project: Part 1
####Author: David Brown

This report will display an exponential distribution. To demonstrate how the Central Limit Theorem works, I will collect the average of 40 exponentials 1000 times. The sample mean will be an unbiased estimate of the population mean, and the sample distribution will be Gaussian.

Using rexp I generated 1000 exponentials with lambda = 0.2. The average of this distribution is roughly 1/0.2; additionally, the standard deviation is also roughly 1/0.2. To demonstrate the Central Limit Theorem, I sampled 40 exponentials from the population 1000 times. I used a for loop and the sample function.

```{r}
set.seed(100)

exp <- rexp(1000, 0.2)
```

```{r}
mns = NULL

for (i in 1:1000) mns = c(mns, mean(sample(exp, 40, replace = TRUE)))
```

```{r, echo=FALSE}
library(ggplot2)
ggplot() + geom_histogram(aes(exp), color = "white", binwidth = 1.5) + geom_vline(xintercept = c(median(exp), mean(exp))) + ggtitle("Exp Distribution")
```

```{r, echo=FALSE}
ggplot() + geom_histogram(aes(mns), color = "white", binwidth = .2) + geom_vline(xintercept = mean(mns), color = "red") + geom_vline(xintercept = 1/0.2, color = "blue") + ggtitle("Distribution of Averages compared to Theoretical Average")
```

Averages
```{r, echo=FALSE}
c(theoretical=1/0.2, pop_average = mean(exp),sample_average = mean(mns))
```
Variances
```{r, echo=FALSE}
c(theoretical=((1/0.2)^2)/1000, pop_distro = var(exp), sample_var = var(exp)/1000)
```
The sample mean is a good, unbiased estimator of the population mean. The population variance is the same as the population mean. The sample distribution is Gaussian.



ggplot() + geom_histogram(aes(exp), color = "white", binwidth = 1.5) + geom_vline(xintercept = c(median(exp), mean(exp))) + ggtitle("Exp Distribution")

ggplot() + geom_histogram(aes(mns), color = "white", binwidth = .2) + geom_vline(xintercept = mean(mns), color = "red") + geom_vline(xintercept = 1/0.2, color = "blue") + ggtitle("Distribution of Averages vs Theoretical Distribution")