---
output: pdf_document
---
#Motor Trend Car Road Tests: Automatic MPG vs Manual MPG
####Executive Summary
The goal of this analysis is to specifically identify which type of transmission (automatic vs manual) is better for MPG. Using the dataset mtcars, a model that isolates the variables which predict mpg will be developed. This report details the exploratory analysis phase, generating various models, testing model variation for model selection, conducting model diagnostics and examining the variability of the outcome. Using this approach, my analysis shows that automatic transmission produces a lower mpg than manual transmission. The analysis shows that manual transmission has almost a 10% higher mpg on average however this estimation has a wide confidence interval.
```{r, echo=FALSE}
##load necessary packages
library(ggplot2)
library(car)
```

```{r, echo=FALSE}
##prep data
dat <- mtcars
dat$am <- as.factor(dat$am)
levels(dat$am) <- c("automatic", "manual")
dat$vs <- as.factor(dat$vs)
```
##Exploratory Analysis
```{r}
##boxplot
ggplot(dat, aes(am, mpg)) + 
  geom_boxplot(fill = "light blue") + 
  ggtitle("Transmission: Auto[0] vs Manual[1]") + xlab("transmission")
```
Boxplot shows Manual Transmission on average has higher MPG.

##Model Selection
```{r, echo=FALSE}
##generate models
fit2.1 <- glm((mpg*10) ~ am + wt -1, "poisson", dat)
fit2.2 <- update(fit2.1, (mpg*10) ~ am + wt + qsec - 1)
fit2.3 <- update(fit2.1, (mpg*10) ~ am + wt + qsec + carb - 1)
fit2.4 <- update(fit2.1, (mpg*10) ~ am + wt + qsec + carb + drat - 1)
fit2.5 <- update(fit2.1, (mpg*10) ~ am + wt + qsec + carb + drat + gear - 1)
##evaluate variance; test models
anova(fit2.1, fit2.2, fit2.3, fit2.4, fit2.5, test = "Chisq")
##model 2 (fit2.2) wins
```
Anova shows that Model 2 (fit2.2) is the best model as far as minimizing residual deviation and deviance while keeping degrees of freedom higher. Model avoids bias by including necessary variables and avoids overfitting.

##Coefficient Interpetation
On average, manual tramsissions have almost 10% higher mpg than automatic (please note that the confidence interval is wide). Additionally, lower bound and upper bound of the 95% confidence interval is higher.
```{r}
##manual vs automatic
1 / (exp(coef(fit2.2))[1] / exp(coef(fit2.2))[2])
##create confidence interval
exp(confint(fit2.2))
```
##Appendix
```{r}
##pairs plot
pairs(dat)
```
Pairs plot shows a matrix of scatterplots to help visualize correlation across predictors (hp, disp, drat, wt are correlated).
```{r}
##correlation matrix
cor(dat[,c(2:7,10:11)])
```
```{r}
##calculate variance inflation for each variable (except vs and am)
sqrt(vif(glm((mpg*10) ~ . -am, "poisson", dat)))
```
```{r}
##model summary
summary(fit2.2)
```
##Model Diagnostics
```{r}
##model diagnostics includes plot of Residuals vs Fitted
par(mfrow = c(2, 2))
plot(fit2.2)
```
